---
title: "Trials with 0.25 > response times < 2.8"
author: "Alba Morató"
date: "25/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
source('helpers/helpers.R')
source('helpers/settings.R')
data <- read.csv("data/behaviour.csv", sep=";")
```

## Original data

```{r data}
# Add continuous time variable (instead of session):
data$time <- recode(data$session, S1 = 0, S2 = 3, S3 = 6, S4 = 12, S5 = 24) 

# Label data:
label(data$trial) = "Trial number"
label(data$block) = "Block number"
label(data$run) = "Run"
label(data$type) = "Memory or non-memory trial"
label(data$S_Angle) = "Stimulus angle (deg)"
label(data$P_Angle) = "Probe angle (deg)"
label(data$R_Angle) = "Response angle (deg)"
label(data$RT) = "Response time (s)"
label(data$MT) = "MT"
label(data$ts_b) = "Beginning timestamp"
label(data$ts_p) = "Probe timestamp"
label(data$ts_d) = "delay timestamp"
label(data$ts_r) = "Response timestamp"
label(data$ts_m) = "Mask timestamp"
label(data$ts_e) = "End timestamp"
label(data$m_angle) = "m_angle"
label(data$m_clock) = "_clock"
label(data$S_rad) = "Stimulus angle (rad)"
label(data$P_rad) = "Probe angle (rad)"
label(data$R_rad) = "Response angle (rad)"
label(data$prevstim_rad) = "Previous stimulus angle (rad)"
label(data$prevresp_rad) = "Previous response angle (rad)"
label(data$prevprob_rad) = "Previous probe angle(rad)"
label(data$prevmem) = "Previous trial is memory?"
label(data$futurestim_rad) = "Next stimulus angle (rad)"
label(data$futureresp_rad) = "Next response angle (rad)"
label(data$subject) = "Subject"
label(data$session) = "Session"
label(data$group) = 'Group'
label(data$error) = "Error (rad)"
label(data$errorprevstim) = "Previous stim error relative to current stim"
label(data$errorprevresp) = "Previous resp error relative to current stim"
label(data$errorprevprobe) = "Previous probe error relative to current stim"
label(data$diffstim) = "Difference current-previous stim (rad)"
label(data$diffresp) = "Difference current stim-previous resp (rad)"
label(data$difffuture) = "Difference current-next stim (rad)"
label(data$difffutureresp) = "Difference current stim-next resp (rad)"
label(data$diffprevprob) = "Difference current stim-previous probe (rad)"
label(data$time) = "Time"

# Remove variables:
data$m_angle <- NULL
data$m_clock <- NULL

# exclude session 5:
data <- data[data$session != 'S5',]

# time from months to years:
data$time <- data$time/12

# Convert variables to factors:
data$session <- as.factor(data$session)
data$session <- droplevels(data$session)
data$group <- as.factor(data$group)
data$subject <- as.factor(data$subject)

# Relevel factors:
data$group <- relevel(data$group, ref = 'C')

# Keep only memory trials:
data = data[data$type == 1, ]

# time from months to days
# data$time <- data$time*30
```


```{r summary}
groupedData <- data %>%
  group_by(group, session) %>%
  nest() %>%
  mutate(n = map(.x = data, .f = ~ n_distinct(.x$subject)),
         n_trials = map(.x = data, .f = ~ nrow(.x))) %>%
  unnest(c(n, n_trials))

groupedData %>%
  select(c(session, group, n)) %>%
  spread(session, n) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Summary of number of subjects by group and session.',
        align = c('l', rep('c', length(unique(data$session)) + 1)), 
        col.names = c(label(data$group), levels(data$session))) %>%
  kable_styling(position = "float_left",
              bootstrap_options = "striped", full_width = F) %>%
  column_spec(1:(length(unique(data$session)) + 1), "2cm") 


groupedData %>%
  select(c(session, group, n_trials)) %>%
  spread(session, n_trials) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Summary of number of trials by group and session.',
        align = c('l', rep('c', length(unique(data$session)) + 1)),
                col.names = c(label(data$group), levels(data$session))) %>%
  kable_styling(position = "center",
                bootstrap_options = "striped", full_width = F) %>%
  column_spec(1:(length(unique(data$session)) + 1), "2cm") 
```

## Descriptive of the raw data

```{r RawPlots, fig.cap = 'Summary of the main variables', fig.show='hold', out.width = '33%', fig.ncol = 3, fig.align='center'}
summaryHist(data, "S_rad")
summaryHist(data, "P_rad")
summaryHist(data, "R_rad")
summaryHist(data, "error")
summaryHist(data, "RT")
```

Let us check trials with response time > 0.25.

```{r}
data = data[data$RT>0.25,]
data = data[data$RT<2.8,]
```

Here we can see all the subject's error distribution by session (just taking into account those trials). 
```{r fig.cap = 'Error distribution by subject and session', fig.height = 40, fig.width = 8}
ggplot(data, aes(x = error, fill = group)) +
  geom_histogram(aes(y =..density..), bins = 30,
                 position = 'identity', alpha = 0.4) +
  scale_fill_manual(values = as.vector(colors)) +
  scale_color_manual(values = as.vector(colors)) +
  labs(x = Hmisc::label(data$error), fill = Hmisc::label(data$group))+
  facet_wrap(~ subject:session, ncol = 4) 
```


```{r summary2}
groupedData <- data %>%
  group_by(group, session) %>%
  nest() %>%
  mutate(n = map(.x = data, .f = ~ n_distinct(.x$subject)),
         n_trials = map(.x = data, .f = ~ nrow(.x))) %>%
  unnest(c(n, n_trials))


groupedData %>%
  select(c(session, group, n_trials)) %>%
  spread(session, n_trials) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Summary of number of trials by group and session.',
        align = c('l', rep('c', length(unique(data$session)) + 1)),
                col.names = c(label(data$group), levels(data$session))) %>%
  kable_styling(position = "center",
                bootstrap_options = "striped", full_width = F) %>%
  column_spec(1:(length(unique(data$session)) + 1), "2cm") 
```

## Bias and precision

```{r}
groupedSubjSes <- data %>%
  group_by(session, subject) %>%
  nest() %>%
  mutate(JV10e = map(.x = data, .f = ~bays_2009_error(X = .x$R_rad, Tg = .x$S_rad)))%>%
  mutate(precision = map_dbl(JV10e, 1), bias = map_dbl(JV10e, 2), 
         group = substr(subject, 1,1), 
         time = unlist(map(data, function(x) unique(x['time'])))) %>%
  select(-c("JV10e")) 

groupedData <- groupedData %>%
  mutate(JV10e = map(.x = data, .f = ~bays_2009_error(X = .x$R_rad, Tg = .x$S_rad)))%>%
  mutate(precision = map_dbl(JV10e, 1), bias = map_dbl(JV10e, 2)) %>%
  select(-c("JV10e"))

aux <- unlist(lapply(levels(data$session), function(x) setNames(2,x)))
aux0 <- apply(expand.grid(c("precision", "bias"), levels(data$session)), 
              1, paste, collapse=".")

groupedData %>%
  select(group, session, precision, bias) %>%
  as.data.frame() %>%
  reshape(idvar = 'group', timevar = 'session', direction = "wide",
          v.names = c('precision', 'bias')) %>%
  select(c("group", all_of(aux0))) %>%
  kable(col.names = c(Hmisc::label(data$group), 
                      rep(c("Precision", "Bias"), length(levels(data$session)))), 
        row.names = FALSE, booktabs = TRUE, escape = FALSE,
        caption = "Bias and precision by group and session.",
        digits = digits) %>%
  add_header_above(c(" " = 1, aux)) %>%
  kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
```

```{r fig.cap = 'Bias and precision by group and session.', fig.show='hold', out.width = '49%'}
groupedSubjSes %>%
  ggplot(aes(x = time, y = abs(bias), color = group, group = group)) +
  geom_point(position= position_jitterdodge(jitter.width = 0.01, dodge.width = 0.05), alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot", show.legend = FALSE,
               position = position_dodge(width=0.05)) +
  stat_summary(fun = mean, geom = "line", size = 1, show.legend = FALSE,
               position=position_dodge(width=0.05)) +
    scale_color_manual(values=colors) +
  labs(x = Hmisc::label(data$time), y = "Absolute bias", 
       color = Hmisc::label(data$group), group = Hmisc::label(data$group))

groupedSubjSes %>%
  ggplot(aes(x = time, y = precision, color = group, group = group)) +
  geom_point(position= position_jitterdodge(jitter.width = 0.01, dodge.width = 0.05), alpha = 0.2) +
  stat_summary(fun.data = "mean_cl_boot",
               position=position_dodge(width=0.05)) +
  stat_summary(fun = mean, geom = "line",  size = 1, show.legend = FALSE,
               position=position_dodge(width=0.05)) +
    scale_color_manual(values=colors) +
  theme(legend.position=c(.9, .8)) +
  labs(x = Hmisc::label(data$time), y = "Precision", color = Hmisc::label(data$group), 
       group = Hmisc::label(data$group)) 
```


### Bias
```{r, fig.align='center', fig.height = 3, fig.cap = 'Bias by subject and group through time'}
groupedSubjSes %>%
  ggplot(aes(x = time, y = abs(bias), color = group, group = subject)) +
  geom_line(alpha = 0.3, show.legend = FALSE) +
  stat_summary(fun.data = "mean_cl_boot", show.legend = FALSE,
               position = position_dodge(width = 0.05), aes(group = group)) +
  stat_summary(fun = mean, geom = "line", size = 1, show.legend = FALSE,
               position = position_dodge(width = 0.05), aes(group = group)) +
  scale_color_manual(values = colors) +
  theme_minimal(base_size = 15) +
  labs(x = Hmisc::label(data$time), y = "Absolute bias", color = Hmisc::label(data$group)) +
  facet_wrap(~ group)
```

#### Modelling bias

##### Time as a continuous variable
```{r eval = FALSE}
mod0 <- lme(abs(bias) ~ group * time, random = ~(1+time)|subject, data = groupedSubjSes)
mod1 <- lme(abs(bias) ~ group * time, random = ~(1)|subject, data = groupedSubjSes)
mod2 <- lm(abs(bias) ~ group * session, 
           data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])
mod3 <- lme(abs(bias) ~ group * session, random = ~1|subject, 
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])
mod4 <- lme(abs(bias) ~ group * session, random = ~(1+session)|subject,
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])

mod0$sigma # best with continous
mod1$sigma
summary(mod2)$sigma
mod3$sigma
mod4$sigma # best with categorical (best model)
```


```{r }
tryCatch(
  summary(lme(abs(bias) ~ group * time, random = ~(1+time)|subject, data = groupedSubjSes))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model where the response is bias and regressors are group, time, its interaction and subject as a random factor.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F),
  error = function(i) {
      summary(lm(abs(bias) ~ group * time, data = groupedSubjSes))$coefficients %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear model where the response is bias and regressors are group, time, its interaction and subject as a random factor.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
  })

tryCatch(
  anova(lme(abs(bias) ~ group * time, random = ~(1+group)|subject, data = groupedSubjSes)) %>%
    kable(digits = digits, booktabs = TRUE, escape = FALSE,
          caption = 'Anova of a linear mixed-effects model where the response is bias and regressors are group, time, its interaction and subject as a random factor.') %>%
      kable_styling(latex_options = "HOLD_position", 
                  bootstrap_options = "striped", full_width = F),
  error = function(i) {
      anova(lm(abs(bias) ~ group * time, data = groupedSubjSes)) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Anova of the linear model where the response is bias and regressors are group, time, its interaction and subject as a random factor..') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
  })
```

##### Time as categorical (session)

```{r }
tryCatch(
  summary(lme(abs(bias) ~ group * session, random = ~(1+session)|subject,
              data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),]))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model where the response is bias and regressors are group, session, its interaction and subject as a random factor..') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F),
  error = function(i) {
      summary(lm(abs(bias) ~ group * session, 
          data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),]))$coefficients %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear model where the response is bias and regressors are group, session, its interaction and subject as a random factor.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
  })

tryCatch(
  anova(lme(abs(bias) ~ group * session, random = ~(1+session)|subject,
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])) %>%
    kable(digits = digits, booktabs = TRUE, escape = FALSE,
          caption = 'Anova of a linear mixed-effects model where the response is bias and regressors are group, session, its interaction and subject as a random factor..') %>%
      kable_styling(latex_options = "HOLD_position", 
                  bootstrap_options = "striped", full_width = F),
  error = function(i) {
      anova(lm(abs(bias) ~ group * session, 
               data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Anova of the linear model where the response is bias and regressors are group, time, its interaction and subject as a random factor..') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
  })
```

###### Just for the encephalitis group (so we can take into account the 4 sessions)
```{r }
summary(lme(abs(bias) ~ session, random = ~(1+session)|subject,
              data = groupedSubjSes[groupedSubjSes$group == 'E',]))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model with data for encephalitis group only. The response is bias and regressor are the sessions and a random factor for subject.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)

anova(lme(abs(bias) ~ session, random = ~(1+session)|subject,
            data = groupedSubjSes[groupedSubjSes$group == 'E',])) %>%
    kable(digits = digits, booktabs = TRUE, escape = FALSE,
          caption = 'Anova of a linear mixed-effects model with data for encephalitis group only. The response is bias and regressors are the sessions and a random factor for subject.') %>%
      kable_styling(latex_options = "HOLD_position", 
                  bootstrap_options = "striped", full_width = F)
```

### Precision

```{r, fig.align='center', fig.height = 3, fig.cap = 'Precision by subject and group through time'}
groupedSubjSes %>%
  ggplot(aes(x = time, y = precision, color = group, group = subject)) +
  geom_line(alpha = 0.3, show.legend = FALSE) +
  stat_summary(fun.data = "mean_cl_boot", show.legend = FALSE,
               position = position_dodge(width = 0.05), aes(group = group)) +
  stat_summary(fun = mean, geom = "line", size = 1, show.legend = FALSE,
               position = position_dodge(width = 0.05), aes(group = group)) +
  scale_color_manual(values = colors) +
  theme_minimal(base_size = 15) +
  labs(x = Hmisc::label(data$time), y = "Precision", color =Hmisc:: label(data$group)) +
  facet_wrap(~ group)
```

#### Modelling precision

```{r eval = FALSE}
rm(mod0, mod1, mod2, mod3, mod4)
mod0 <- lme(precision ~ group * time, random = ~(1+time)|subject, data = groupedSubjSes)
mod1 <- lme(precision ~ group * time, random = ~(1)|subject, data = groupedSubjSes)
mod2 <- lm(precision ~ group * session, 
           data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])
mod3 <- lme(precision ~ group * session, random = ~1|subject, 
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])
mod4 <- lme(precision ~ group * session, random = ~(1+session)|subject,
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])

mod0$sigma # best with continuous
mod1$sigma
summary(mod2)$sigma
mod3$sigma
mod4$sigma # best with categorical
```


##### Time as a continuous variable

```{r}
# remove -Inf or Inf values
groupedSubjSes <- groupedSubjSes[!is.infinite(rowSums(groupedSubjSes[,c(4,5)])),]

summary(lme(precision ~ group * time, random = ~(1)|subject, data = groupedSubjSes))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model where the response is precision and regressors are group, time, its interaction and subject as a random factor.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)

anova(lme(precision ~ group * time, random = ~(1)|subject, data = groupedSubjSes)) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = "Anova for the linear mixed-effects model where the response is precision and regressors are group, time, its interaction and subject as a random factor..") %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
```

##### Time as a categorical variable (session)

```{r}
summary(lme(precision ~ group * session, random = ~(1)|subject, 
            data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),]))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model where the response is precision and regressors are group, session, its interaction and subject as a random factor..') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)

anova(lme(precision ~ group * session, random = ~(1)|subject, 
          data = groupedSubjSes[groupedSubjSes$session %in% c('S1', 'S4'),])) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = "Anova for the linear mixed-effects model where the response is precision and regressors are group, session, its interaction and subject as a random factor..") %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
```

###### Just for the encephalitis group (so we can take into account the 4 sessions)
```{r}
summary(lme(precision ~  session, random = ~(1+session)|subject, 
            data = groupedSubjSes[groupedSubjSes$group == 'E',]))$tTable %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = 'Linear mixed-effects model with data for encephalitis group only. The response is the precision and regressors are the sessions and a random factor for subject.') %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)

anova(lme(precision ~  session, random = ~(1+session)|subject, 
          data = groupedSubjSes[groupedSubjSes$group == 'E',])) %>%
  kable(digits = digits, booktabs = TRUE, escape = FALSE,
        caption = "Anova for the linear mixed-effects model with data for encephalitis group only. The response is the precision and regressors are the sessions and a random factor for subject.") %>%
    kable_styling(latex_options = "HOLD_position", 
                bootstrap_options = "striped", full_width = F)
```
